---
title: "Reproducible research - week2.rmd"
author: "Raghu Maan"
date: "3/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible research - week 2 assignment

###1. Code for reading in the dataset and/or processing the data

```{r}
setwd("/Users/t93ku6h/Documents/Development/coursera/Reproducible-Research/week2")
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity_data <- read.csv('activity.csv')
```

####Summary
```{r}
summary(activity_data)
```

###2. Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
steps_per_day <- tapply(activity_data$steps, activity_data$date, sum, na.rm=TRUE)
qplot(steps_per_day, xlab='Steps per day', ylab='Frequency', binwidth=1000)
```

###3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(steps_per_day)
```

###4. Time series plot of the average number of steps taken
```{r}
library(dplyr)
#split the activity data by time internval (0,5,10..) and run the mean function over the splitted data
mean_steps_per_interval <- aggregate(x=list(steps=activity_data$steps), by=list(interval=activity_data$interval), FUN=mean, na.rm=TRUE)
head(mean_steps_per_interval)
#plot interval by steps
with(mean_steps_per_interval,plot(interval,steps,type="l",ylab="average number of steps",xlab="5-minute time interval"))
```

###5. The 5-minute interval that, on average, contains the maximum number of steps
mean_steps_per_interval[which.max(mean_steps_per_interval$steps),"interval"]

###6. Code to describe and show a strategy for imputing missing data 
```{r}
#install.packages("mice")
library(mice)
#show missing data - shows 2304 missing entries
md.pattern(activity_data,plot=TRUE)
#impute data with pmm method
#https://statisticalhorizons.com/predictive-mean-matching
activity_data_imputed <- mice(activity_data,m=1,method = "pmm")
```

###7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
activity_data_imputed_completed <- complete(activity_data_imputed,1)
steps_per_day <- tapply(activity_data_imputed_completed$steps, activity_data_imputed_completed$date, sum)
qplot(steps_per_day, xlab='Steps per day', ylab='Frequency', binwidth=1000)
```

###8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
activity_data$day <-  ifelse(as.POSIXlt(activity_data$date)$wday %in% c(0,6), 'weekend', 'weekday')
activity_data_with_dayofweek <- aggregate(steps ~ interval + day, data=activity_data, mean)
ggplot(activity_data_with_dayofweek, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(day ~ .) +
    xlab("5-minute time interval") + 
    ylab("avarage number of steps")
```